---
title: "Fortnite_Weapons_Type_Analysis"
author: "vaspoon"
date: "2023-09-16"
output:
  pdf_document: default
  html_document: default
---

#Import needed packages

```{r, echo=FALSE}
require(dplyr)
require(ggplot2)
require(viridis)
require(ggcorrplot)
require(FactoMineR)
```

#Preparing Data \## Import Dataset and Cleaning it

```{r}
weapons <- read.csv("data/Fortnite_BR_Weapon_Attributes.csv",sep=";")
weapons$Name <- as.factor(weapons$Name)
weapons$Rarity <- as.factor(weapons$Rarity)
weapons$Type <- as.factor(weapons$Type)
```

## Dataset overview

```{r}
summary(weapons)
```

Critical..
and Crit..Damage are corrupted columns, its values are always equal to 0.
Durability is also a corrupted column as its value is always "?".
We'll be removing these columns for the following analysis.

```{r}
weapons <- select(weapons, -c("Critical..","Crit..Damage","Durability"))
```

```{r}
cat("Distinct Weapons :",length(levels(weapons$Name)),"\n",
"Distinct Rarity Levels :",length(levels(weapons$Rarity)),"\n",
"Distinct Weapon Types :",length(levels(weapons$Type)))
```

This dataset contains 16 different weapons overall.
Weapons are from 6 different Types, and 5 disitinct levels of Rarity are are presented.

# Exploring Weapons Types & Rarity

## Barchart per Type

```{r}
weapons_type_barchart <- ggplot(data = weapons,aes(x=Type,fill = Rarity)) + geom_bar(stats="identity",alpha = 0.8,color="black") +theme_minimal() +theme(axis.text.x = element_text(angle=90))+ xlab("Weapon Type") +
  ylab("Number of Weapons")

weapons_type_barchart <- weapons_type_barchart + scale_fill_manual(values=c('#B2B2B2','#DE1FF6','#FFA500','#11BFE1','#7EC091')) #Colors usually used in video games for rarity levels

weapons_type_barchart
```

## Damages per Weapon Type

Looking at DPS repartition among different weapon types.

```{r}
dps_type_violin <- ggplot(weapons, aes(x = Type, y = DPS))+   geom_violin(aes(fill = Type), trim = FALSE) +geom_boxplot(width = 0.1)+ theme_minimal() +theme(axis.text.x = element_text(angle=90))+ xlab("Weapon Type") + ylab("DPS")

dps_type_violin <- dps_type_violin + scale_fill_manual(values=c('#B78181','#7583b0','#FFA500','#a1bb93','#b589a3','#b0a275')) # Palette color manually scaled for Weapon Types

dps_type_violin
```

It seems like Pistols is the weapon type with the wider range of DPS.
Weapon types with the most different DPS distributions are Sniper Riffles and Submachine Guns.

## Barchart per Rarity

```{r}

weapons_rarity_barchart <- ggplot(data = weapons,aes(x=Rarity,fill = Type )) + geom_bar(alpha = 0.8,color="black") +theme_minimal() +theme(axis.text.x = element_text(angle=90)) + xlab("Rarity Category") + ylab("Number of Weapons")

weapons_rarity_barchart <- weapons_rarity_barchart + scale_fill_manual(values=c('#B78181','#7583b0','#FFA500','#a1bb93','#b589a3','#b0a275'))

weapons_rarity_barchart
```

## Damages per Rarity

```{r}
dps_rarity_violin <- ggplot(weapons, aes(x = Rarity , y = DPS))+  geom_violin(aes(fill = Rarity), trim = FALSE)+ geom_boxplot(width = 0.1) +theme_minimal() +theme(axis.text.x = element_text(angle=90)) + xlab("Weapon Type") + ylab("DPS")

dps_rarity_violin <- dps_rarity_violin + scale_fill_manual(values=c('#B2B2B2','#DE1FF6','#FFA500','#11BFE1','#7EC091')) # Palette color manually scaled for Weapon 

dps_rarity_violin

```

It seems more difficult to differentiate Rarity Levels of weapons studying the distributions of DPS among weapons.

# Characterizing Weapons with PCA

## Correlations Matrix

First we'll temporarely keep only numeric variables

```{r}
temp.weapons <- select(weapons, -c(1,10,11))
corr <- round(cor(temp.weapons), 1)
ggcorrplot(corr, method = "circle" , type = "lower")
```

## PCA

### Applying PCA

```{r}

weapons_PCA <- select(weapons,-c(1))

res.PCA <- PCA(X = weapons_PCA,quali.sup = c(9,10),graph = F)

cat("First Dimension explains",res.PCA$eig[1,2],"% of the dataset's Variance","\n","Second Dimension explains",res.PCA$eig[2,2],"% of the dataset's Variance","\n\n","Meaning",res.PCA$eig[1,2]+res.PCA$eig[2,2],"% is explained in the first two dimensions")
```

### Studying Dimensions

```{r}
dimdesc(res.PCA,axes = c(1,2))
```

-   Dimension 1

Is positively correlated to Mag..Size, Fire.Rate and DPS.
Is negatively correlated to Impact and Damage

-   Dimension 2

Is positively correlated to Reload.Time and Range

Both dimensions describe most part of information we found with the correlation Matrix.

### Variables

```{r}
plot.PCA(res.PCA, axes=c(1, 2), choix="var", habillage="cos2")
```

Following the dimensions' description, the graph of variables offers a better representation of variables' correlations with dimensions.

Moreover,by observing the cos² of each variables, we notice that the Ammo.Cost variable is not properly projected on the factorial plan.
Therefore we will not make any conclusion regarding this variable.

Also, it is reassuring that the 1st Dimension presents opposite correlations for Fire.Rate and Damage variables, as this is a well known trade-off in game design for weapons.

### Individuals

We're now going to analyze weapons Rarity Levels and Types representations in the factorial plan.

```{r}
plotellipses(res.PCA)
```

-   Weapon Rarity Levels

As confidence ellipses of each Rarity level overlaps with each of them, it is not possible to differentiate weapons of each category on the factorial plan.

This is in fact a great thing as it means each Rarity level doesn't differentiate from another in a significant way regarding for example Damages,DPS and Impact which could possibly impact the balance between weapons.

-   Weapon Types

The first thing we can notice is that the 2nd Dimension differentiates Shotguns from all other weapon types, while the 1st Dimension presents a more homogeneous distribution of weapon types.
Here is some conclusions we can find for all weapon types :

-   Shotguns are mostly defined by higher hit Range and Reload Time values than other types, but also by high values of Damage and Impact.

-   Shotguns are mostly defined by higher hit Range and Reload Time values than other types.

-   Submachine Guns and Assault Rifles are defined by higher DPS,Fire Rate and MagSize, and lower values in Impact and Damage.

-   Sniper Riffles and Explosives are defined by higher values in Damages and Imapct, and lower values in DPS, Fire Rate and MagSize.


